---
authors:
 - 'Pelin PEKER'
 - 'Merve AK'
 - 'Edanur Binnaz DURSUN'
 - 'Ahmet ÇALI'
date: 'Mayıs 2024'
institution: 'FEN FAKÜLTESİ'
division: 'İSTATİSTİK BÖLÜMÜ'
advisor: 'Dr. Engin YILDIZTEPE'
department: 'İstatistik Bölümü'
degree: 'Bitirme Projesi Raporu'
title: 'DEĞİŞİM NOKTASI'
knit: 'bookdown::render_book'
site: 'bookdown::bookdown_site'

output: 
  thesisdown::thesis_pdf: default #"template.tex" şablonuna yönlendiren fonksiyon
#  thesisdown::thesis_gitbook: default
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default
# If you are creating a PDF you'll need to write your preliminary content here or
# use code similar to line 20 for the files.  If you are producing in a different
# format than PDF, you can delete or ignore lines 20-31 in this YAML header.

abstractTR: |
  `r if(knitr:::is_latex_output()) paste(readLines("00-ozet.Rmd", encoding="UTF-8"), collapse = '\n  ')`

abstract: |
  `r if(knitr:::is_latex_output()) paste(readLines("00-abstract.Rmd"), collapse = '\n  ')`
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab is
# needed on the line after the |.

preface: |  # rapor değerlendirme
  "`r rmarkdown::metadata$title`" başlıklı bitirme projesi raporu tarafıgitmdan okunmuş, kapsamı ve niteliği açısından bir Bitirme Projesi raporu olarak kabul edilmiştir.\
  \
  \
  `r rmarkdown::metadata$advisor`
  
acknowledgements: |
  Tüm çalışma süresince yönlendiriciliği, katkıları ve yardımları ile yanımızda olan danışmanımız `r rmarkdown::metadata$advisor` 'ye ve böyle bir çalışmayı yapmamız için bize fırsat tanıyan Dokuz Eylül Üniversitesi Fen Fakültesi İstatistik Bölümüne teşekkür ederiz.\
  \
  \
  `r rmarkdown::metadata$authors[[1]]`
  \
  `r rmarkdown::metadata$authors[[2]]`
  \
  `r rmarkdown::metadata$authors[[3]]`
  \
  `r rmarkdown::metadata$authors[[4]]`
  \
 
# ```{r, echo = FALSE,results="asis"}
# n<-length(rmarkdown::metadata$authors)
# for (i in 1:n) {
#       cat(rmarkdown::metadata$authors[[i]],sep="\n")
#   }
# ```
#

#dedication: |
#  You can have a dedication here if you wish. 
bibliography: bib/thesis.bib
# Download your specific bibliography database file and refer to it in the line above.
csl: csl/apa-tr.csl
# Download your specific csl file and refer to it in the line above.
lot: true
lof: true
#space_between_paragraphs: true
# Delete the # at the beginning of the previous line if you'd like
# to have a blank new line between each paragraph
#header-includes:
#- \usepackage{tikz}
---

#  thesisdown::thesis_gitbook: default

Placeholder



<!--chapter:end:index.Rmd-->


# Değişim Noktası {#Bolum2}

Placeholder


## Tek Değişim Noktası Tespiti
## Birden Fazla Değişim Noktası Tespiti
### İkili Segmentasyon Algoritması
### PROPHET
### PELT(Pruned Exact Linear Time)
#### Optimal Bölütleme
#### PELT Yöntemi
## Parçalı Regresyon
## KAPSAMA METRİĞİ
## F1 PUANI
#### Doğruluk (Accuracy)
#### Duyarlılık (Recall)
#### Kesinlik (Precision)
#### F1 Skoru (F1 Score)

<!--chapter:end:01-chap1.Rmd-->


# Uygulama {#Bolum3}

Placeholder


## Veri
### Gerçek Veri 
### Yapay Veri
### F1 Default
### Cover Default
### F1 Oracle
### Cover Oracle
## Yapay Veriler
### F1 Default
### COVER Default
### F1 Oracle
### COVER Oracle
## Sporcu Verileri
### Cover Oracle
### F1 Oracle
## Çalışma Tasarımı
## Sonuçlar

<!--chapter:end:02-chap2.Rmd-->

# Bölüm Başlığı {#Bolum4}
## Bu bir alt başlık

Bu bölümde şu konular yer almaktadır...

### Bu ikinci seviye bir alt başlık

<!--chapter:end:03-chap3.Rmd-->

# Bölüm 4 Başlık {#Bolum5}

## Bu bir alt başlık

Bu bölümde şu konular yer almaktadır...

### Bu ikinci seviye bir alt başlık

<!--chapter:end:04-chap4.Rmd-->

# Sonuç {-}

If we don't want Conclusion to have a chapter number next to it, we can add the `{-}` attribute.

**More info**

And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph. Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.


<!--chapter:end:77-conclusion.Rmd-->


# Kaynaklar {-}

Placeholder



<!--chapter:end:88-references.Rmd-->

`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'` 

<!--
If you feel it necessary to include an appendix, it goes here.
-->


# İlk Ek Başlığı

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.

**In the main Rmd file**

```{r ref.label='include_packages', results='hide', echo = TRUE}
```

**In Chapter \@ref(ref-labels):**

```{r ref.label='include_packages_2', results='hide', echo = TRUE}
```

# İkinci Ek Başlığı

İkinci Ek

<!--chapter:end:99-appendix.Rmd-->

